\subsection{Basic Increasing and Maintaining Rules}
 
The construction of the application conditions introduced in the previous section, aswell as the constructed application conditions itself, are very complex.
For a certain set of rules, which we will call \emph{basic consistency increasing rules},
for which, we are able to construct application conditions with the same property,
namely that a rule equipped with this application condition is consistency increasing at layer, in a less complex manner.  
The main idea is, that these rules are (a) not able to delete occurrences of existentially or insert occurrences of universally bound graphs and (b) are able to increase consistency at a certain layer. 
That means, given a basic increasing rule $\rho$, there does exist a transformation $t: G \Longrightarrow_{\rho} H$ such that $t$ is a consistency increasing transformation w.r.t to a constraint $c$.

To ensure that (a) is met, we firstly introduce \emph{basic consistency maintaining rules up to layer}, which means that, given a constraint, a plain rule is not able 
to delete existentially bound and insert universally bound graphs up to a certain layer. 
For the definition, we use the notion of consistency maintaining rules up to layer. 
The set of basic consistency maintaining rules up to layer is indeed a subset of the 
set of consistency maintaining rules up to layer since these rules have to be plain rules, whereas consistency maintaining rules up to layer are allowed to be 
equipped with application conditions, i.e. $\maintaining(\cdot,\cdot)$.


\begin{definition}[\textbf{basic consistency maintaining rule up to layer}]
	Let a plain rule $\rho$ and a constraint $c$ in UANF be given.
	Then, $\rho$ is called \emph{basic consistency maintaining rule up to layer
	$-1 \leq k < \nlvl(c)$ w.r.t. $c$} if $\rho$ is a direct consistency maintaining 
	rule at layer $k$.
\end{definition}

\begin{example}
	Consider the rules \emph{\texttt{moveFeature}} and 
	\emph{\texttt{assignFeature}} given in Figure \ref{fig:rules} and constraint
	$c_1$ given in Figure \ref{fig:constraints}.
	The rule \emph{\texttt{assignFeature}} is a basic consistency maintaining 
	rule up to layer $1$ w.r.t. $c$, whereas \emph{\texttt{moveFeature}} is 
	not a basic consistency maintaining rule.
\end{example}

Since infinite many transformations via a plain rule $\rho$ exist, it is impossible to check whether $\rho$ is a basic consistency maintaining rule up to layer based on the definition above. Therefore, we present a characterisation of basic consistency maintaining rules, which only relies on $\rho$ itself. 
First, let us assume that $\rho$ is able to create occurrences of a universally bound graph $C_j$. 
This is possible if (a) $\rho$ does insert an edge of $C_j \setminus C_{j-1}$ which connects already existing nodes of $C_j$, since it is unclear whether this would create a new occurrence of $C_j$, or (b) if $\rho$ does insert a node $v$ of $C_j$ such that all edges $e \in E_{C_j}$ with $\src(e) = v$ or $\tar(e) = v$ are also inserted. 
If at least one of these edges is not inserted, it is guaranteed that this insertion does not create an occurrence of $C_j$ since $v$ is only connected to edges that have also been inserted by $\rho$.

Second, let us assume that $\rho$ is able to delete occurrences of a existentially bound graph $C_j$.
This is possible if (a) $\rho$ does delete an edge of $C_j \setminus C_{j-1}$ or (b) $\rho$ deletes a node $v$ of $C_j \setminus C_{j-1}$ such that all edges $e \in E_{C_j}$ with $\src(e) = v$ or $\tar(e) = v$ are also deleted. 
If $\rho$ deletes a node $c$ of $C_{j} \setminus C_{j-1}$ without all its connected edges in $C_j$ there does not exist a transformation via $\rho$ such 
that an occurrence of $C_j$ is deleted by deleting this node since the dangling edge condition would not be satisfied. 
A rule satisfying this properties does not decrease the satisfaction at layer. 
Additionally, we have to ensure that the number of violations will not be 
increased. For this, we have to check that $\rho$ is not able to insert 
occurrences of the corresponding universally bound, as described above, and
that $\rho$ is not able to remove occurrence of any intermediate graph. 
This is only ensured, if $\rho$ does not remove any elements of $C'$ if the 
set of intermediate graphs is given by $\ig{C}{C'}$.

To check that a plain rule satisfies these properties, we make use of the dangling edge condition, or in other words, we check that the rule is not applicable at 
certain overlaps of $L$ and an existentially bound graph or that the inverse
rule is not applicable at certain overlaps of $R$ and an universally bound graph.

\begin{lemma}\label{def:non-decreasing}
	Let a plain rule $\rho = L \xhookleftarrow{l} K \xhookrightarrow{r} R$ and a 
	constraint $c$ in EANF be given. Then, $\rho$ is a basic consistency 
	maintaining rule up to layer $-1 \leq k < \nlvl(c)$ w.r.t $c$ if 
	\ref{category:non-decreasing_1} and \ref{category:non-decreasing_2} apply 
	for all $k$  and \ref{category:non-decreasing_3} applies if $k \leq \nlvl(c) 
	-3$ and $\scond{k}{c}$ is existentially bound, i.e. $k$ is odd.  
	\begin{enumerate}
		\item \label{category:non-decreasing_1}
			Let $E$ be the set of all existentially bound graphs $C_j$ with
			$0 \leq j \leq k+1$.
			For each graph $C \in E$ and each overlap $P \in \overlay(L,C)$ 
			with $i_L^P(L \setminus K) \cap i_C^P(C) \neq \emptyset$, 
			the transformation
			$$t: P \Longrightarrow_{\rho, i_L^P} H$$ does not exist.


		\item \label{category:non-decreasing_2}
			Let $U$ be the set of all universally bound graphs $C_j$ with
			$0 \leq j \leq k+2$.
			For each graph $C \in U$ and each overlap $P \in \overlay(L,C)$ 
			with $i_R^P(R \setminus K) \cap i_C^P(C) \neq \emptyset$ 
			the transformation
			$$t: P \Longrightarrow_{\rho^{-1}, i_R^P} H$$ does not exist.		  

		\item \label{category:non-decreasing_3}
			For all graphs $P \in \overlay(L, C_{k+3})$ it holds that
			$$i_L^P(L \setminus K) \cap i_{C_{k+3}}^P(C_{k+3} \setminus C_{k+1})
			= \emptyset
			$$
\end{enumerate}

\end{lemma}
\begin{proof}
	Let $\rho = L \xhookleftarrow{l} K \xhookrightarrow{r} R$ be a rule 
	that satisfies the characterisations listed in Lemma 
	\ref{def:non-decreasing} with $-1 \leq k < \nlvl(c)$. Let us assume that 
	$\rho$ is not a direct consistency maintaining rule up to layer $k$ w.r.t. 
	$c$.
	Then, there exists a transformation $t: G \Longrightarrow_{\rho,m} H$ with 
	$\maxk{c}{G} = k$ such that $t$ is not direct consistency maintaining w.r.t. 
	$c$.
	Therefore, either (\ref{direct_improving_1}), (\ref{direct_improving_1_2}),
	(\ref{direct_improving_3}) or (\ref{direct_improving_4}) is not satisfied.
	\begin{enumerate}
		\item 
			If (\ref{direct_improving_1}) is not satisfied, then $k < \nlvl(c) 
			-2$, let $j = \maxk{c}{G}
			+2$, there does exist an occurrence $p:C_j \inj G$ such that 
			$p \models \ic{0}{\scond{j}{c}}{C}$ and $\track_t \circ p \models 
			\ic{0}{\scond{j}{c}}{C}$ with $C \in \ig{C_j}{C_{j+1}}$.
			Since $i_L^P(L \setminus K) \cap i_{C_{k+3}}^P(C_{k+3} \setminus 
			C_{k+1})= \emptyset$ for all $P \in \overlay(L, C_{k+3})$ this 
			case can to apply since $\rho$ does not delete any elements 
			of $C_{j+1}\setminus C_j$.
			
		\item 
			If  (\ref{direct_improving_1_2}) is not satisfied, there does 
			exist an occurrence $p:C_j \inj H$ such that no morphism 
			$q: C_j \inj G$ with $p = \track_t \circ q$ exists and 
			and $p \not \models \false$ if $\scond{j}{c} = \false$ and 
			$p \not \models \ic{0}{\scond{j}{c}}{C_j+1}$ otherwise.
			Since $\rho$ satisfies \ref{category:non-decreasing_2}
			$G$ must have dangling edges and therefore, $G$ is not a graph.
			
		\item 
			If (\ref{direct_improving_3}) is not satisfied, there does 
			exist an occurrence $p : C_j \inj H$ of an universally bound graph 
			with $j \leq \maxk{c}{G}$ such that no morphism $q:C_j \inj G$ with 
			$\track_t \circ q = p$ exists. 
			Again, since $\rho$ satisfies \ref{category:non-decreasing_2}
			$G$ must have dangling edges and therefore, $G$ is not a graph.
		\item
			If (\ref{direct_improving_4}) is not satisfied, there does 
			exist an occurrence $p : C_j \inj H$ of an existentially bound graph 
			with $j \leq \maxk{c}{G}$ such that $\track_t \circ p$ is not total.
			Since $\rho$ satisfied \ref{category:non-decreasing_1}
			the dangling edge condition would not bee satisfied and $H$ would 
			contain dangling edges.
	\end{enumerate}	    
	In total follows that $\rho$ is a direct basic consistency maintaining rule 
	up to 
	layer $k$.
\end{proof}

Now, we are ready to introduce \emph{basic increasing rules at layer $k$} with $k$ being odd. The set of basic increasing rules is a subset of the set of maintaining rules at layer $k$ which ensures that the satisfaction at layer as well as the number of violation will not be decreased.
Additionally, the left-hand side of these rule do contain an occurrence $p$ of the  universally bound graph $C_{k+2}$, such that this occurrence either will be 
removed,  i.e. elements of $C_{k+2}\setminus C_{k+1}$ will be deleted, or an intermediate graph $C \in \ig{C_{k+2}}{C_{k+3}}$ will be inserted.
Of course, this second case only occurs if $k < \nlvl(c) -3$ with $c$ being the respective constraint. 
This property yields the advantage that application conditions for basic increasing rules are less complex and smaller, since it can be exactly determined how this rule removes a violation  and therefore, no overlaps have to be considered.

This, on first sight seems like a restriction of the set of basic increasing rules but the context of each rule $\rho$ that does satisfy all properties of a basic increasing rule excluding that $C_{k+2}$ is a sub-graph of the left-hand side can be expanded such that $\rho$ is a basic increasing rule and the semantic of $\rho$ is not increased. 
Later on, a method to derive this rules will be presented.

Basic increasing rules at layer $k$ are called \emph{deleting basic increasing rules} if $p$ will be removed  and \emph{inserting basic increasing rules} if $\scond{k}{c}$ an intermediate graph will be inserted.
For our repairing process, we will introduce the restriction that deleting basic increasing rules are only allowed to delete edges but no nodes of $C_{k+2}$ since otherwise it is not possible, given a rule set and a constraint to decide whether this rules set is able to repair an arbitrary graph based only on deleting basic increasing rules. 
For example, consider a rule that deletes a node of $C_{k+2}$.
Then, it is unknown whether this node is connected to nodes not belonging to $C_{k+2}$ and it is unclear whether all occurrence of $C_{k+2}$ could be destroyed by $\rho$ since the dangling edge condition might be unsatisfied. 


\begin{definition}[\textbf{basic increasing rule}]\label{def_basicIncreasing}
	Let a constraint $c$ in UANF and a direct consistency maintaining rule 
	$\rho = (\ac,  L \xhookleftarrow{l} K \xhookrightarrow{r} R)$ up to layer 
	$-1 \leq k < \nlvl(v) -1$, with $k$ odd, be given. Then, $\rho$ is called  
	\emph{basic increasing w.r.t $c$ at layer 
	$k$} if a morphism $p: C_{k+2} 
	\inj L$, called the \emph{increasing morphism}, exists such that either 
	\ref{incr_1} or \ref{incr_2} applies.
	
	\begin{enumerate}
		\item \label{incr_1}
			$r \circ l^{-1} \circ p$ is not total. Then, $\rho$ is called
			a \emph{deleting basic increasing rule}.
		\item \label{incr_2}	
			If $k < \nlvl(c)-2$, there does exist an intermediate graph
			$C \in \ig{C_{k+2}}{C_{k+3}}$ such that $p \not \models \exists(
			a_{k+2}^r:C_{k+2} \inj C, \true)$, $r \circ l^{-1} \circ p$ is 
			total and 
			$r \circ l^{-1} \circ p \models \exists(a_{k+2}^r:C_{k+2} \inj C,
			\true)$. Then, $\rho$ is called a \emph{inserting basic increasing 
			rule with $C$}. 
	\end{enumerate} 
\end{definition}

\begin{example}
	Consider the rule \emph{\texttt{assignFeature}} given in Figure 
	\ref{fig:rules} and constraint $c_1$ given in Figure \ref{fig:constraints}.
	Then, assign Feature is a inserting basic rule with $C_2^2 \in \ig{C_1^1}
	{C_2^1}$ w.r.t. $c_1$ but not a inserting basic rule with respect to the 
	constraint $\forall(C_2^2, \exists C_2^1)$ since the left-hand side 
	of  \emph{\texttt{assignFeature}} does not contain an occurrence 
	of $C_2^2$.
\end{example}
Again, \ref{def_basicIncreasing} relies on every transformation of a rule $\rho$. 
Therefore, we present an alternative method to determine whether $\rho$ satisfies \ref{def_basicIncreasing} or not by checking that $\rho$ does not delete any edges or nodes of $C_{k+1} \setminus C_k$. 

As mentioned above, given a consistency maintaining rule $\rho$ we can derive 
basic increasing rules that are only applicable if at a match and graph if 
$\rho$ is applicable.


\begin{definition}[\textbf{derived rules}]
	Let a constraint $c$ in UANF and an rule $\rho = (\ac,  L \xhookleftarrow{l} 
	K \xhookrightarrow{r} R)$ be given. 
	The set of \emph{derived rules of $\rho$} at layer $ 0 \leq k \leq 
	\nlvl(c)-1$ 
	contains rules characterized in the following way:
	Let 
	$$\mathbf{G} := \begin{cases}
						\ig{C_k}{C_{k+1}} & \text{if $\scond{k}{c}$ is 
						existentially bound} \\
						\{C_k\} &\text{otherwise}
					\end{cases}
	$$
	For each $P \in \mathbf{G}$ and $L' \in \overlay(L,P)$: If the diagram shown 
	in Figure 
	\ref{fig_dpo_construction} is a transformation, i.e. (1) and (2) are 
	pushouts, and for the characterisations of Definition 
	\ref{def_basicIncreasing} holds that
	\begin{equation*}
		\begin{split}
			&\rho \text{ satisfies \ref{incr_1}} \implies \rle{L'}{l'}{K'}{r'}{R'} 
			\text{ satisfies \ref{incr_1} } \wedge \\
			&\rho \text{ satisfies \ref{incr_2}} \implies \rle{L'}{l'}{K'}{r'}{R'} 
			\text{ satisfies \ref{incr_2}}
		\end{split}
	\end{equation*}	 
	the rule $$\rho' = (\shiftm(\ac, i_L^{L'}), \rle{L'}{l'}{K'}{r'}{R'})$$
	is a derived rule of $\rho$ at layer $k$.
\end{definition}
\input{figures/fig_dpo_construction_basic}

\begin{example}
	Consider the rule \emph{\texttt{assignFeature}} given in Figure 
	\ref{fig:rules} and constraint $c_1$ given in Figure \ref{fig:constraints}.
	The set of derived rules of $\rho$ at layer $1$ is given in Figure 
	\ref{fig:derived}.
\end{example}

\input{figures/fig_derived_rules}

Obviously, a rule $\rho'$ contained in the set of derived rules of a rule $\rho$ is only
applicable at a match $m'$ if $\rho$ is applicable at match $m$ with 
$m = m' \circ i$ with $i$ being the inclusion of the left-hand side of $\rho$ into 
the left-hand side of $\rho'$.
Therefore, given a rules set $\mathcal{R}$, the extension of $\mathcal{R}$ by the set of all derived rules for each rule of $\mathcal{R}$ does not extend the expressiveness of $\mathcal{R}$.
The main idea of the concept of derived rules is the extension of a given rule set by as many basic increasing rules as possible without extending the expressiveness of this set.  

\begin{lemma}
	Let a constraint $c$ in UANF, and a rule $\rho = (\ac, \rle{L}{l}{K}{r}{R})$ 
	be given, such that $\rho$ is 
	a maintaining rule up to layer $-1 \leq k < \nlvl(c)-1$, with $k$ odd, and
	satisfies \ref{incr_1} and \ref{incr_2} of Definition 
	\ref{def_basicIncreasing}. Then, each rule contained in the set of derived 
	rules of $\rho$ at layer $k$ is a basic increasing rule. 
\end{lemma}

\begin{proof}
	Let $\rho' = (\ac', \rle{L'}{l'}{K'}{r'}{R'})$ be one of those derived rules.
	Since $\rho'$ deletes an inserts exactly the same elements as $\rho$ 
	and $m' \models \ac' \iff m' \circ i_L^{L'} \models$, $\rho'$ is a 
	consistency maintaining rule up to layer $k$ and satisfies \ref{incr_1} and 
	\ref{incr_2} of Definition \ref{def_basicIncreasing}.
	Since a morphism $i: C_k \inj L$ exists, $\rho'$ is a basic increasing 
	rule at layer $k$. 
\end{proof}

In transformations via a rule $\rho$, such that the match intersects with an occurrence of an universally bound graph $C_k$, $\rho$ can be replaced by a derived rule of $\rho$ at layer $k$.  

\begin{lemma}
	Let a constraint $c$ in UANF and a rule $\rho = (\ac, \rle{L}{l}{K}{r}{R})$ 
	be given. 
	Then, for each transformation
		$$t: G \Longrightarrow_{\rho,m}H$$ 
	such that an occurrence $p:C_k \inj G$ of a universally bound graph $C_k$ 
	with $p(C_k) \cap m(L) \neq \emptyset$ exists, there does exist a 
	transformation
		$$t': G \Longrightarrow_{\rho', m'}H$$
	with $\rho'$ being a derived rule of $\rho$ at layer $k$.
\end{lemma}

\begin{proof}
	Since $p(C_k) \cap m(L) \neq \emptyset$ there does exist an overlap
	$P \in \overlay(C_k,L)$ such that a morphism $q: \inj G$ with 
	$m = q \circ i_L^P$ and $p = q \circ i_{C_k}^P$ exists. 
	Since $t$ exists, there does exist a derived rule $\rho' = (\ac', 
	\rle{L'}{l'}{K'}{r'}{R})$ with $L' = P$.
	We set $m' = p$, since $m = m' \circ i_L{L'} \models \ac$ it follows 
	that $m' = \ac'$.
	Since $\rho$ does removes and inserts the same elements as $\rho$,
	there does exist a transformation
	$t': G \Longrightarrow_{\rho',m'} H$.  
\end{proof}

Via this, we are able to replace consistency increasing transformations
via a rule $\rho$ that is consistency maintaining up to layer $k$ and satisfies satisfies \ref{incr_1} and \ref{incr_2} of Definition \ref{def_basicIncreasing}
by a derived rule of $\rho$ at layer $k$, that is, a basic increasing rule at 
layer $k$. 

%\begin{lemma}\label{lemma:equiv_direct_normal}
%Let a constraint $c$ in UANF, a basic increasing rule $\rho$ at layer $\kmax +1$ and a transformation $t: G \Longrightarrow_{\rho} H$ be given. 
%Then, 
%$$t \text{ is consistency increasing w.r.t $c$} \iff t \text{ is direct consistency increasing w.r.t $c$}$$
%\end{lemma}
%\begin{proof}
%The \enquote{$\Longleftarrow$} side of the equivalence holds with lemma \ref{lemma:direct_implies_normal}. 
%We need to show that the \enquote{$\Longrightarrow$} side of the equivalence holds. 
%Let $t: G \Longrightarrow_{\rho} H$ be a consistency increasing transformation and $\rho$ a basic increasing rule. 
%Then, $t$ satisfies (\ref{def_basicIncreasing}) and the satisfaction of (\ref{direct_improving_1}) follows immediately.
%Since $\rho$ is also a non-consistency decreasing rule up to layer $\kmax$, $t$ satisfies (\ref{non_consistency_universally}) and (\ref{non_consistency_existentially}), then the satisfaction (\ref{direct_improving_1_2}), (\ref{direct_improving_3}) and (\ref{direct_improving_4}) also follows immediately.
%\begin{enumerate}
%\item If $\rho$ is a deleting basic increasing rule. 
%Since $t$ is a consistency increasing rule there has to exist a morphism $p:C_{\kmax+2} \inj G$ such that $\track_t \circ p$ is not total and with that (\ref{direct_improving_2}) is satisfied. 
%\item If $\rho$ is a inserting basic increasing rule there exists an occurrence $p: C_{\kmax + 2} \inj G$ and a graph $C \in \ig{C_{\kmax +2}}{C_{\kmax+3}}$ such that $p \not \models \ic{\kmax +2}{c}{C}$ and $\track_t \circ p \not \models \ic{\kmax +2}{c}{C}$. 
%Therefore (\ref{direct_improving_2}) is satisfied. 
%\end{enumerate}
%In total follows that $t$ is a direct consistency increasing transformation w.r.t $c$.
%\end{proof}


%\input{Kapitel/basic improving rules/comparison_basic_rules}



